### 2022-05-26 main/chapter_2ljets.tex
\,\eV over ~\eV; minor figure caption improvement

\begin{figure}[t]
\centering
\includegraphics[width=0.48\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
\hfill
\includegraphics[width=0.45\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
\caption{%
Supersymmetric signal processes in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Left: C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
weak bosons in decays are bounded by the mass splitting
$m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
We explore the parameters
$m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
\\[0.5em]
Right: GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
is produced by soft decays from pairs including $\chargino_1$,
$\neutralino_2$ or $\neutralino_1$.
Although the $h$ and $Z$ bosons decay to many states, we target
$Z\rightarrow \ell\ell$ with
$h/Z\rightarrow bb/jj$.
We explore the parameters
$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with fixed
$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1~\eV[G]$.
}
\label{fig:2ljets_signal_diagrams}
\end{figure}

@@

\clearpage
\FloatBarrier
\section{Context}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run 2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13~\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

\subsubsection{Former excesses}
Excess data were observed in several signal regions of the partial Run 2
search ``for chargino--neutralino production using recursive
jigsaw reconstruction in final states with two or three charged
leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
$\sqrt{s} = 13~\eV[T]$.
That search reported notable excesses in four signal regions.
Of these, two required two charged leptons
($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
and other two required three
($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).

@@

In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26~\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
\begin{itemize}
\item Electron pair: requirements vary between
\begin{itemize}
\item $\pt^{e_2} > 12~\eV[G]$ and
\item $\pt^{e_2} > 17~\eV[G]$.
\end{itemize}
\item Muon pair: various requirements include
\begin{itemize}
\item $\pt^{\mu_2} > 10~\eV[G]$,
\item $\pt^{\mu_2} > 14~\eV[G]$,
\item $\pt^{\mu_1} > 18~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$, and
\item $\pt^{\mu_1} > 22~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$.
\end{itemize}
\item Emu pair: various asymmetric requirements include
\begin{itemize}
\item $\pt^{e_1} > 17~\eV[G]~\&~\pt^{\mu_1} > 14~\eV[G]$,
\item $\pt^{e_1} > 7~\eV[G]~\&~\pt^{\mu_1} > 24~\eV[G]$, and
\item $\pt^{e_1} > 26~\eV[G]~\&~\pt^{\mu_1} > 8~\eV[G]$.
\end{itemize}
\end{itemize}
Quite generally throughout the data used in $\twoljets$, trigger requirements
have increased in hardness with time.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger do not also harmfully increase inseparable backgrounds.
A study, \textit{which was performed by Knut Vadla}, investigated the effects
of of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200~\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $\approx10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We considered that to be too little to be worthwhile.

@@

Although jet and $\met$ triggers were not studied in detail, they can be
quickly ruled out since they hard objects: examples of these triggers require
approximately $\ptjone > 400~\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200~\eV[G]$, or
$\met > 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
Leptons, however, are striking signals which are produced as smaller background
rates, as they suppressed by electroweak couplings.
Therefore, \atlas\ can trigger on leptons with much softer requirements;
lepton triggers allow us to use softer jets and $\met$, which are relied upon
throughout the $\twoljets$ analyses.

@@

Baseline leptons are required only to have  $\pt > 3~\eV[G]$ for muons and
$\pt > 4.5~\eV[G]$ for electrons.
Signal leptons are all required to have $\pt > 10~\eV[G]$, along with tighter
quality and isolation requirements, details of which are given in the
paper~\cite{atlas2022searches}.
Electrons are accepted within $|\eta| < 2.47$.
Baseline muons have $|\eta| < 2.7$, and signal muons have $|\eta| < 2.6$.

All leptons used in regions of the $\twoljets$ electroweak analysis are signal
leptons with $\pt > 25~\eV[G]$.
This harder requirement reduces backgrounds, moves our selections away from the
lowest trigger thresholds, and matches the precedent set by previous
work~\cite{atlas_23l_SUSY_2016_24}.

@@

Baseline jets are defined to require $\pt > 20~\eV[G]$ with $|\eta| < 2.8$;
signal jets additionally require $\pt > 30~\eV[G]$.
To reduce the impact of pile-up jets, a variety of quality criteria are used,
depending on the jet $\pt$ and $|\eta|$.
Details of these quality criteria and angular selections are given in the
paper~\cite{atlas2022searches}.

@@

Signal $b$-tagged jets require $\pt > 20~\eV[G]$, $|\eta| < 2.5$ and a tag from
the `MV2c10' algorithm at its $77\%$ efficiency working point.
\TODO{cite, summarize}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
example signal models.
At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
signal leptons with
$\mll \in (71, 111)~\eV[G]$,
$\mjj \in (60, 110)~\eV[G]$,
$\nbtag \leq 1$,
$\met > 100~\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
standard model processes; dileptonic $t\bar t$ and $W^\pm W^\mp$ decays both
satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
The $\twoljets$ electroweak search depends on this effect;
most selections require $\mttwo(\ell^+, \ell^-, 0) > 80~\eV[G]$, or similar,
to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.48\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
\hfill
\includegraphics[width=0.45\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
\caption{%
Supersymmetric signal processes in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Left: C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
weak bosons in decays are bounded by the mass splitting
$m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
We explore the parameters
$m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
\\[0.5em]
Right: GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
is produced by soft decays from pairs including $\chargino_1$,
$\neutralino_2$ or $\neutralino_1$.
Although the $h$ and $Z$ bosons decay to many states, we target
$Z\rightarrow \ell\ell$ with
$h/Z\rightarrow bb/jj$.
We explore the parameters
$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with fixed
$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1\,\eV[G]$.
}
\label{fig:2ljets_signal_diagrams}
\end{figure}

@@

\clearpage
\FloatBarrier
\section{Context}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run 2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

\subsubsection{Former excesses}
Excess data were observed in several signal regions of the partial Run 2
search ``for chargino--neutralino production using recursive
jigsaw reconstruction in final states with two or three charged
leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$.
That search reported notable excesses in four signal regions.
Of these, two required two charged leptons
($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
and other two required three
($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).

@@

In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26\,\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
\begin{itemize}
\item Electron pair: requirements vary between
\begin{itemize}
\item $\pt^{e_2} > 12\,\eV[G]$ and
\item $\pt^{e_2} > 17\,\eV[G]$.
\end{itemize}
\item Muon pair: various requirements include
\begin{itemize}
\item $\pt^{\mu_2} > 10\,\eV[G]$,
\item $\pt^{\mu_2} > 14\,\eV[G]$,
\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
\end{itemize}
\item Emu pair: various asymmetric requirements include
\begin{itemize}
\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
\end{itemize}
\end{itemize}
Quite generally throughout the data used in $\twoljets$, trigger requirements
have increased in hardness with time.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger do not also harmfully increase inseparable backgrounds.
A study, \textit{which was performed by Knut Vadla}, investigated the effects
of of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $\approx10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We considered that to be too little to be worthwhile.

@@

Although jet and $\met$ triggers were not studied in detail, they can be
quickly ruled out since they hard objects: examples of these triggers require
approximately $\ptjone > 400\,\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200\,\eV[G]$, or
$\met > 100\,\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
Leptons, however, are striking signals which are produced as smaller background
rates, as they suppressed by electroweak couplings.
Therefore, \atlas\ can trigger on leptons with much softer requirements;
lepton triggers allow us to use softer jets and $\met$, which are relied upon
throughout the $\twoljets$ analyses.

@@

Baseline leptons are required only to have  $\pt > 3\,\eV[G]$ for muons and
$\pt > 4.5\,\eV[G]$ for electrons.
Signal leptons are all required to have $\pt > 10\,\eV[G]$, along with tighter
quality and isolation requirements, details of which are given in the
paper~\cite{atlas2022searches}.
Electrons are accepted within $|\eta| < 2.47$.
Baseline muons have $|\eta| < 2.7$, and signal muons have $|\eta| < 2.6$.

All leptons used in regions of the $\twoljets$ electroweak analysis are signal
leptons with $\pt > 25\,\eV[G]$.
This harder requirement reduces backgrounds, moves our selections away from the
lowest trigger thresholds, and matches the precedent set by previous
work~\cite{atlas_23l_SUSY_2016_24}.

@@

Baseline jets are defined to require $\pt > 20\,\eV[G]$ with $|\eta| < 2.8$;
signal jets additionally require $\pt > 30\,\eV[G]$.
To reduce the impact of pile-up jets, a variety of quality criteria are used,
depending on the jet $\pt$ and $|\eta|$.
Details of these quality criteria and angular selections are given in the
paper~\cite{atlas2022searches}.

@@

Signal $b$-tagged jets require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a tag from
the `MV2c10' algorithm at its $77\%$ efficiency working point.
\TODO{cite, summarize}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
example signal models.
At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This selection requires two same-flavor opposite-sign signal leptons with
$\pt^{\ell_2} > 25\,\eV[G]$,
$\mll \in (71, 111)\,\eV[G]$,
$\mjj \in (60, 110)\,\eV[G]$,
$\nbtag \leq 1$,
$\met > 100\,\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
standard model processes; dileptonic $t\bar t$ and $W^\pm W^\mp$ decays both
satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
The $\twoljets$ electroweak search depends on this effect;
most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
